---
title: "HW3"
author: "Ali Harb"
date: "February 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(pracma)
library(MASS)
```

## 1. Problem set 1

#### (1) What is the rank of the matrix A?
$$\mathbf{A} = \left[\begin{array}
{rrr}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3
\end{array}\right]
$$

```{r}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3),nrow=4)
A
```

Reduce to echelon form and find the rank

```{r}
rref(A)
Rank(A)
```

The Rank of A is 4

#### (2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

+ If m > n then  the maximum rank of the matrix would n Rank.
+ If the matrix has non-zero elemnt then the minimum ramk would be 1.

#### (3) What is the rank of matrix B?
$$\mathbf{B} = \left[\begin{array}
{rrr}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2
\end{array}\right]
$$

By looking at matrix B we can easily determine that it has a rank of 1 because row 2 and 3 are scalar multiples of row 1. Thus, they are linearly dependent.

Get the rank using the Rank function we will get the same result

```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2),nrow=3)
B
```

Reduce to echelon form and find the rank

```{r}
rref(B)
Rank(B)
```

## 2. Problem set 2

#### Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$$\mathbf{A} = \left[\begin{array}
{rrr}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6
\end{array}\right]
$$

#### Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.

Solution

Find the eigenvalue of A using the rule of sarrus:

$$\mathbf{p(\lambda)} = \left[\begin{array}
{rrr}
\lambda -1 & -2 & -3 & \lambda -1 & -2 \\
0 & \lambda -4 & -5 & 0 & \lambda -4 \\
0 & 0 & \lambda -6 & 0 & 0
\end{array}\right]
$$

$$ p(\lambda) = (\lambda -1)(\lambda-4)(\lambda-6)+0+0-0-0-0 $$ 

$$ (\lambda -1)= 0 => \lambda = 1$$
$$ (\lambda -4)= 0 => \lambda = 4$$
$$ (\lambda -6)= 0 => \lambda = 6$$

We can also use the eigen function to find the eigen values of A matrix 

```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6),nrow=3, byrow=T)
A

Lambda<-eigen(A)$value
Lambda

```

Find the eigenvectors of A:

### Solve for $\lambda = 1$

```{r}
L=1
L1 <- matrix(c(L-1,-2,-3,0,L-4,-5,0,0,L-6),nrow=3, byrow=T)
L1
```

Reduce to echelon form

```{r}
rref(L1)
```

$( \lambda I_3 - A )v = \left[ {\begin{array}{ccc} 
0 & 1 & 0\\
0 & 0 & 1\\
0 & 0 & 0\ \end{array} } 
\right] \left[ {\begin{array}{c}
v_1\\ 
v_2\\ 
v_3\ \end{array} } \right]$

$v_2 = 0 \\ v_3 = 0 \\ v_1 = 1$

Let $v_1 = a$

$E_{\lambda = 1} = a \left[ {\begin{array}{c} 1\\ 0\\ 0\ \end{array} } \right]$


### Solve for $\lambda = 4$

```{r}
L=4
L4 <- matrix(c(L-1,-2,-3,0,L-4,-5,0,0,L-6),nrow=3, byrow=T)
L4
```

Reduce to echelon form

```{r}
r=rref(L4)
fractions(r)
```

$( \lambda I_3 - A )v = \left[ {\begin{array}{ccc} 
1 &-2/3 & 0\\
0 & 0 & 1\\
0 & 0 & 0\ \end{array} } 
\right] \left[ {\begin{array}{c}
v_1\\ 
v_2\\ 
v_3\ \end{array} } \right]$

$v_2 = 2/3 v_2 \\ v_3 = 0 \\ v2 = 1$

Let $v_2 = b$

$E_{\lambda = 4} = b \left[ {\begin{array}{c} 2/3\\ 1\\ 0\ \end{array} } \right]$


### Solve for $\lambda = 6$

```{r}
L=6
L6 <- matrix(c(L-1,-2,-3,0,L-4,-5,0,0,L-6),nrow=3,byrow=T)
L6
```

Reduce to echelon form

```{r}
r=rref(L6)
fractions(r)
```


$( \lambda I_3 - A )v = \left[ {\begin{array}{ccc} 
1 & 0 & -8/5\\
0 & 1 & -5/2\\
0 & 0 & 0\ \end{array} } 
\right] \left[ {\begin{array}{c}
v_1\\ 
v_2\\ 
v_3\ \end{array} } \right]$

$v_1 = 8/5 v_3 \\ v_2 = 5/2 v_3 , v3=1$

Let $v_3 = c$

$E_{\lambda = 6} = c \left[ {\begin{array}{c} 8/5\\ 5/2\\ 1\ \end{array} } \right]$










