---
title: "Assignment 3"

output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assinment Task

Is to study the dataset and the associated description of the data (i.e. "data dictionary"). You may need to look around a bit, but it's there! You should take the data, and create a data frame with a subset of the columns (and if you like rows) in the dataset. You should include the column that indicates edible or poisonous and three or four other columns. You should also add meaningful column names and replace the abbreviations used in the data-for example, in the appropriate column, "e" might become "edible." Your deliverable is the R code to perform these transformation tasks. 

## Get the data

Reading the data form the CSV file using the provided URL link. Extracting
the data and store them into my_data  

```{r cars}
require(RCurl)

myurl <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
e_data <- getURL(myurl)

my_data <- read.csv(text=e_data,header=FALSE,sep=",")
head(my_data)
```

## Construct the data dictionary 

Creating a subset with desireed fields from the stored data    

```{r}
my_subset <- my_data[,c(1,4,6,9,20,23)]
knitr::kable(head(my_subset))
```

## Rename the fields 

Giving the column a meaningful name that represent the data   

```{r}
colnames(my_subset) <-c("class","cap-color","odor","gill-size","ring-type","habitat")

knitr::kable(head(my_subset))
```

## Replace abbreviation 

Use the levels function to replace the abbreviated characters with it's attribute categories 

```{r}
# column class 
levels(my_subset$'class') <- c(levels(my_subset$'class'), "edible", "poison")
my_subset$'class'[my_subset$'class' == "e"] <- "edible"
my_subset$'class'[my_subset$'class' == "p"] <- "poison"

# column cap-color 
levels(my_subset$'cap-color') <- c(levels(my_subset$'cap-color'),
                                   "brown", "buff","cinnamon","gray","green","pink","purple","red","white","yellow","bruises","n0")
my_subset$'cap-color'[my_subset$'cap-color' == "n"] <- "brown"
my_subset$'cap-color'[my_subset$'cap-color' == "b"] <- "buff"
my_subset$'cap-color'[my_subset$'cap-color' == "c"] <- "cinnamon"
my_subset$'cap-color'[my_subset$'cap-color' == "g"] <- "gray"
my_subset$'cap-color'[my_subset$'cap-color' == "r"] <- "green"
my_subset$'cap-color'[my_subset$'cap-color' == "p"] <- "pink"
my_subset$'cap-color'[my_subset$'cap-color' == "u"] <- "purple"
my_subset$'cap-color'[my_subset$'cap-color' == "e"] <- "red"
my_subset$'cap-color'[my_subset$'cap-color' == "w"] <- "white"
my_subset$'cap-color'[my_subset$'cap-color' == "y"] <- "yellow"
my_subset$'cap-color'[my_subset$'cap-color' == "t"] <- "bruises"
my_subset$'cap-color'[my_subset$'cap-color' == "f"] <- "n0"

# column odor
levels(my_subset$'odor') <- c(levels(my_subset$'odor'),
                              "almond", "anise","creosote", "fishy","foul","musty","none","pungent","spicy")
my_subset$'odor'[my_subset$'odor' == "a"] <- "almond"
my_subset$'odor'[my_subset$'odor' == "l"] <- "anise"
my_subset$'odor'[my_subset$'odor' == "c"] <- "creosote"
my_subset$'odor'[my_subset$'odor' == "y"] <- "fishy"
my_subset$'odor'[my_subset$'odor' == "f"] <- "foul"
my_subset$'odor'[my_subset$'odor' == "m"] <- "musty"
my_subset$'odor'[my_subset$'odor' == "n"] <- "none"
my_subset$'odor'[my_subset$'odor' == "p"] <- "pungent"
my_subset$'odor'[my_subset$'odor' == "s"] <- "spicy"

# column gill-size
levels(my_subset$'gill-size') <- c(levels(my_subset$'gill-size'), "broad", "narrow")
my_subset$'gill-size'[my_subset$'gill-size' == "b"] <- "broad"
my_subset$'gill-size'[my_subset$'gill-size' == "n"] <- "narrow"

# column ring-type
levels(my_subset$'ring-type') <- c(levels(my_subset$'ring-type'),
                                   "cobwebby", "evanescent","flaring", "large","none","pendant","sheathing","zone")
                             
my_subset$'ring-type'[my_subset$'ring-type' == "c"] <- "cobwebby"
my_subset$'ring-type'[my_subset$'ring-type' == "e"] <- "evanescent"
my_subset$'ring-type'[my_subset$'ring-type' == "f"] <- "flaring"
my_subset$'ring-type'[my_subset$'ring-type' == "l"] <- "large"
my_subset$'ring-type'[my_subset$'ring-type' == "n"] <- "none"
my_subset$'ring-type'[my_subset$'ring-type' == "p"] <- "pendant"
my_subset$'ring-type'[my_subset$'ring-type' == "s"] <- "sheathing"
my_subset$'ring-type'[my_subset$'ring-type' == "z"] <- "zone"

# column habitat
levels(my_subset$'habitat') <- c(levels(my_subset$'habitat'),
                              "grasses", "leaves","meadows", "paths","urban","waste","woods")
my_subset$'habitat'[my_subset$'habitat' == "g"] <- "grasses"
my_subset$'habitat'[my_subset$'habitat' == "l"] <- "leaves"
my_subset$'habitat'[my_subset$'habitat' == "m"] <- "meadows"
my_subset$'habitat'[my_subset$'habitat' == "p"] <- "paths"
my_subset$'habitat'[my_subset$'habitat' == "u"] <- "urban"
my_subset$'habitat'[my_subset$'habitat' == "w"] <- "waste"
my_subset$'habitat'[my_subset$'habitat' == "d"] <- "woods"

knitr::kable(head(my_subset))

```

