---
title: "HW Week 5"
author: "Ali Harb"
date: "October 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
The objective of this assignment is to practice tidying and transforming data using R studio or MySQL.  

## Software Requirement
To perform the task various libraries are needed for data transformation and visualization.


```{r message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)

```

## Loading Data
A csv file of the data must be created and imported into R for analytical processing.

```{r}

flight_timedelay= data.frame( 
  airlines = c('alaska', 'alaska','am_West','am_West'),
  status = c('on_time', 'delayed', 'on_time','delayed'),
  losAngeles = c(497,62,694,117),
  phoenix = c(221,12,4840,415),
  sanDiego = c(212,20,383,65),
  sanFrancisco = c(503,102,320,129),
  seattle = c(1841,305,201,61)
)

write.table(flight_timedelay, file = "flight_timedelay.csv", sep = ",", col.names=T)

airlines_tb <-data.frame(read.csv("https://raw.githubusercontent.com/aliharb/R-Programming/master/flight_timedelay.csv"))

airlines_tb
str(airlines_tb)

```

## Data Analysis 

Based on the data provided, analysis will be performed to find the following:

* The Best airline performance 
* Airport with least delay

1. transform the data to group based on delays and ontime arrival.

```{r}

delays_ontime<-filter(airlines_tb,status=="on_time")
ontimes_tb<-gather(delays_ontime,'airports','flights',3:7)

cols_delay<-filter(airlines_tb,status=="delayed")
delay_tb<-gather(cols_delay,'airports','flights',3:7)

tb<-bind_rows(ontimes_tb,delay_tb)
tb
```

Plot on time arrivals for both airlines 

```{r}
ggplot(ontimes_tb,aes(x=airports, y=flights)) +
  geom_bar(stat="identity",fill="green", colour="black") +
  facet_wrap(~ airlines) +
  labs(title="Comparison of Arrival On Time", x="Airports", y="On time")
```

Plot delays arrivals for both airlines  

```{r}
ggplot(delay_tb,aes(x=airports, y=flights)) +
  geom_bar(stat="identity",fill="red", colour="black") +
  facet_wrap(~ airlines) +
  labs(title="Comparison of Arrival Delayed", x="airports", y="Delayed")
```

From the above graphs we can determine the following:

* Am west has more airline traffic 
* Phoenix airport has the highest arrival flights 

Even though the graphs highlight important information about the data set but they didn't show any evidence about airlines and airport performance  

2. at first, the proportionality and percentile of ever evry flights arrivals will compared.

```{r set-options, echo=FALSE, cache=FALSE}
options(width=100)
delays_ontime<-spread(tb,status,flights)
delays_ontime<-mutate(delays_ontime,total_flights=delayed+on_time)

delays_ontime<-mutate(delays_ontime,delay_proportional=delayed/total_flights)
delays_ontime<-mutate(delays_ontime,delay_percentile=delay_proportional*100)
delays_ontime

airport_performance<-delays_ontime %>%
group_by(airports) %>% 
summarise(airport_delays = sum(delayed), airport_flights=sum(total_flights),airport_proportional=(airport_delays/airport_flights)*100)
```


```{r}
ggplot(airport_performance,aes(x=airports, y=airport_proportional)) +
  geom_bar(stat="identity",fill="blue", colour="black") +
  labs(title="Comparison of Airpot performance", x="airports", y="percentile")
```

The graph indicates that, Phoenix airport has the best performance with the lowest delay less than 8% delays.

```{r}
airlines_status<-tb%>%group_by(airlines, status) %>% 
  summarise(flights = sum(flights))
airlines_performance<-spread(airlines_status,status,flights)
airlines_performance<-mutate(airlines_performance,Proportion = delayed / (on_time+delayed)) 
airlines_performance<-mutate(airlines_performance,percentile = Proportion*100) 
airlines_performance
```

```{r}
ggplot(airlines_performance,aes(x=airlines, y=percentile)) +
  geom_bar(stat="identity",fill="yellow", colour="black") +
  labs(title="Comparison of Airlines Delayed", x="airlines", y="percentile")
```


Based on the graph, the am west airline has a better performance with less than 11% delays.

















