---
title: "Project4"
author: "Ali Harb"
date: "November 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this project information from flights database will be migrated from MySQL(relational database) to MongoDB, which is NoSQL  .

## Libraries 

Loading the required libraries to perform the task using R studio. 
```{r message=FALSE, warning=FALSE}
library(knitr)
library(RMySQL)
library(RSQLite)
library(rmongodb)
library(mongolite)
```

## Load Data

Load the flights dataset from MySQL to R environment. 

Set password as mypassword:
```{r echo=FALSE}
mypassword <-'alex'
```

Get data from MySQL.

```{r}

mydrv <- dbDriver("MySQL")
mydb <- dbConnect(mydrv, user ='root', password = mypassword, dbname='flights')

dbListTables(mydb)
dbListFields(mydb, 'flights')

mydata <- dbGetQuery(mydb,"SELECT month, day, 
                     dep_time, arr_time, air_time,  
                     dep_delay, arr_delay, distance,
                     carrier FROM flights")

kable(head(mydata))
kable(tail(mydata))
```

Observe the data structure of the imported object.

```{r}
str(mydata)
```

Substitute NA with 0, since all Na are integers data type.

```{r}
mydata[is.na(mydata)]<-0
kable(tail(mydata))
```

Disconnect the MySQL database.

```{r}
dbDisconnect(mydb)
```

## Export Data

Export the R data frame to mongodb, the NoSQL database.

Establish connection with mongodb server:

```{r}
mongodata = mongo.create(host = "localhost")
mongo.is.connected(mongodata)
```

Create the database and then upload it into mongo.

```{r}
# create the database in mongodb
mongodata <- mongo(collection="documents",db="flights")

# put column header into a list
feilds<-colnames(mydata)

#establish connection with mongodb
mongodata = mongo.create(host = "localhost")
mongo.is.connected(mongodata)

# remove all record in database if it exist
for(i in 1:length(feilds)){
  st<- paste("flights",feilds[i], sep=".")
  print(mongo.remove(mongodata, ns=st))
}

# Check if records are removed 
for(i in 1:length(feilds)){
  DBNS<-paste("flights", feilds[i], sep=".")
  print(mongo.count(mongodata,ns=DBNS))
}

# insert the records to mongodb flights database
for(i in 1:length(mydata)){
  tmp<-as.list(mydata[i])
  bson=mongo.bson.from.list(tmp)
  st<- paste("flights", feilds[i], sep=".")
  mongo.insert(mongodata,st, bson)
}
```

Check if the dataset is loaded to mongodb server by reloading the dataset to R. 

```{r}
# if databse exists
mongo.get.databases(mongodata)

# create dataframe and a list
df<-data.frame(stringsAsFactors = FALSE)
lst<-list()

# get the data into a list
for(i in 1:9){
  st<- paste("flights", feilds[i], sep=".")
  tmplst<-mongo.find.all(mongodata, ns= st, mongo.oid2character = T)
  tmplst[[1]][1]<-NULL
  lst<-c(lst,tmplst)
}
#convert list to datframe
df<-as.data.frame(lst,stringsAsFactors = FALSE)

str(df)

# display the data 
kable(head(df))

```

## Advantages of NoSQL over SQL

NoSQL can handle large volumes data, unstructured dataset, allow multiple data structures, and object oriented programming.  in addition to these create benefits it is also efficient and scalable. When working in an evolving project, NoSQL will provide a more particle option; otherwise, SQL is more practical for structural and logical related data with a standard and proven technology.


## References  

[rmongodb][https://gist.github.com/Btib7751989ert3/]

[rmongo][https://www.r-bloggers.com/rmongo-accessing-mongodb-in-r/]


